---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Overview

mergen employs artificial intelligence to convert data analysis questions into executable code, explanations, and algorithms. The self-correction feature ensures the generated code is optimized for performance and accuracy. mergen features a user-friendly chat interface, enabling users to interact with the AI agent and extract valuable insights from their data effortlessly.

### Installation
 ```{r,eval=FALSE}
 # the easiest way to install mergen is via install.packages
 install.packages("mergen")
 ```

#### Development version
 
To get a bug fix or to use a feature from the development version, you can install the development version of mergen from GitHub.
 
 ```{r,eval=FALSE}
 devtools::install_github()
 ```
 
### Prerequisites

- Make an AI account
- Create an AI API key to use with the package
- Set up the API key in R

#### Configuring your AI API key

To interact with an AI API, you require a valid AI API key. To configure your key so that it is present globally in your environment at all times, you can include it in your .Renviron file. This will ensure that the key is automatically loaded. 

For setting up mergen with openai, this variable should be called `OPENAI_API_KEY`. 
If you are planning on using another API, call this variable `AI_API_KEY`. For more information on setting up an agent, we recommend you visit __Get Started__ . 

__Caution:__ If you’re using version control systems like GitHub, include .Renviron in your .gitignore file to prevent exposing your personal API key. 

Here is how to open your .Renviron file for modification in your project:

```{r,eval=FALSE}
require(usethis)

edit_r_environ(scope="project")
```

For a persistent loading of your API key add one of the following lines to your .Renviron file based on the agent you plan to use:

```{r,eva=FALSE}
# setting up for openAI agents:
OPENAI_API_KEY = "your_key"

# setting up for other agents:
AI_API_KEY = "your_key"
```

If you however wish to set this variable only for a single session, you can use on of the the following command, based on the agent you plan to use:

```{r,eval=FALSE}
# setting up for openAI agents:
Sys.setenv(OPENAI_API_KEY = "your_key")

# setting up for other agents:
Sys.setenv(AI_API_KEY = "your_key")

```
 
### Getting help
 
If you encounter a clear bug, please file an issue with a minimal reproducible example on GitHub. There you can also post further questions. 

